{\bf Affine.m} -- {\it Mathematica} package for Lie algebras of finite, affine and extended affine type

### 2011-02-16 AffineLieAlgebras/CPC/article.tex
CPC paper started

\title{Recursive properties of branching and BGG resolution}
\author{V D Lyakhovsky$^1$ and A A Nazarov$^{2,3}$\\
{\small $^1$ Theoretical Department, SPb State University}\\
{\small 198904, Sankt-Petersburg, Russia}\\
{\small e-mail: lyakh1507@nm.ru}\\
{\small $^2$ Theoretical Department, SPb State University}\\
{\small 198904, Sankt-Petersburg, Russia}\\
{\small$^{3}$ Chebyshev Laboratory,}\\
{\small Department of Mathematics and Mechanics, SPb State University}\\
{\small 199178, Saint-Petersburg, Russia}\\
{\small email: antonnaz@gmail.com}}

@@

\begin{abstract}
Recurrent relations for branching coefficients are based on a special type
of singular element decomposition. We show that this decomposition can be
used to construct the parabolic Verma modules and finally to obtain the
generalized Weyl-Verma formulas for characters. We demonstrate how branching coefficients can determine the generalized BGG resolution sequence.
\end{abstract}

\section{Introduction}

\label{sec:introduction}

Branching properties of Lie (affine Lie) algebras are highly important for
applications in quantum field theory (see for example the conformal field
theory models \cite{difrancesco1997cft},\cite{coquereaux2008conformal}). In
this paper we demonstrate that for an arbitrary reductive subalgebra
branching is directly connected with the BGG resolution and in particular
exhibits the resolution properties in terms of the $\mathcal{O}^{p}$
category \cite{lepowsky1977generalization} (the parabolic generalization of
the cathegory $\mathcal{O}$ \cite{bernstein1976category}).

The resolution of irreducible modules in terms of infinite-dimensional ones is important for the
theory of integrable spin chains \cite{derk1008}. In the Baxter $\mathcal{Q}$-operator approach \cite{derk09} the generic transfer
matrices corresponding to the (generalized) Verma modules are factorized into the product of Baxter
operators. The resolution allows to calculate the transfer matrices for finite-dimensional
auxiliary spaces.

To show the connection of the BGG resolution with the branching we use the recursive approach presented in \cite
{2010arXiv1007.0318L} (similar to the one used in \cite{ilyin812pbc} for
maximal embeddings). We consider the subalgebra $\af \hookrightarrow \gf$ together with its
counterpart $\afb$ ''orthogonal'' to $\af$ with respect to
the Killing form and also $\widetilde{\afb}:=\afb\oplus \frak{h}_{\perp }$ where $\frak{h}=\frak{\frak{h}_{\af}}\oplus
\frak{h}_{\afb}\oplus \frak{h}_{\perp }$. For any reductive
subalgebra $\af$ the subalgebra $\afb\hookrightarrow \gf$
is regular and reductive. For a highest weight integrable module $%
L^{\left(\mu \right) }$ and orthogonal subalgebra $\af_{\bot }$ we
consider the singular element $\Psi ^{\left( \mu \right) }$ (the numerator
in the Weyl character formula $ch\left( L^{\mu }\right) =\frac{\Psi ^{\left(
\mu \right) }}{\Psi ^{\left( 0\right) }}$, see for example \cite
{humphreys1997introduction}) and the Weyl denominator $\Psi _{\af_{\bot
}}^{\left( 0\right) }$ for the orthogonal partner. It is shown that the element $\Psi _{\gf%
}^{\left( \mu \right) }$ can be decomposed into a combination of Weyl
numerators $\Psi _{\af_{\bot }}^{\left( \nu \right) }$ with $\nu \in P_{%
\mathfrak{a}_{\bot}}^{+}$. This decomposition provides the possibility to
construct the set of highest weight modules $L_{\afb}%
^{\mu _{\afb}}$. When the injection \ $\af%
_{\bot }\hookrightarrow \gf$ \ satisfies the ''standard parabolic''
conditions these modules give rise to the parabolic Verma modules $M_{\left(
\afb \hookrightarrow \gf\right) }^{\mu _{%
\afb}}$ so that the initial character $ch\left(
L^{\mu }\right) $ is finally decomposed into the alternating sum of such. On
the other hand when the parabolic conditions are violated the construction
survives and exhibits a decomposition with respect to a set of generalized Verma
modules $M_{\left( \widetilde{\frak{b}_{\perp }},\gf\right) }^{\mu _{%
\widetilde{\afb}}}$ where $%
\widetilde{\frak{b}_{\perp }}$ is not a subalgebra in $\gf$ but a
contraction of $\widetilde{\afb}$.

Some general properties of the proposed decompositions are formulated in
terms of a specific formal element $\Gamma _{\af\rightarrow \gf}$  called ''the injection fan''. Using this tool a simple and explicit algorithm for branching
rules applicable for an arbitrary (maximal or
nonmaximal) subalgebra in affine Lie algebras was
proposed in \cite{2010arXiv1007.0318L}.

Possible further developments are discussed in Section \ref{sec:conclusions}.

\subsection{Notation}

\label{sec:notation}

Consider Lie algebras (affine Lie algebras) $\frak{g}$ and $\frak{a}$ and an
injection $\frak{a}\hookrightarrow \frak{g}$ such that $\frak{a}$ is a
reductive subalgebra $\frak{a\subset g}$ with correlated root spaces: $\frak{%
h}_{\frak{a}}^{\ast }\subset \frak{h}_{\frak{g}}^{\ast }$. We use the
following notations:

$\frak{g=n}^{-}+\frak{h}+\frak{n}^{+}$ --- the Cartan decomposition;

$r$ , $\left( r_{\frak{a}}\right) $ --- the rank of the algebra $\frak{g}$ $%
\left( \mathrm{{resp. }\frak{a}}\right) $ ;

$\Delta $ $\left( \Delta _{\frak{a}}\right) $--- the root system; $\Delta
^{+} $ $\left( \mathrm{{resp. }\Delta _{\frak{a}}^{+}}\right) $--- the
positive root system (of $\frak{g}$ and $\frak{a}$ respectively);

$\mathrm{mult}\left( \alpha \right) $ $\left( \mathrm{mult}_{\frak{a}}\left(
\alpha \right) \right) $ --- the multiplicity of the root $\alpha$ in $%
\Delta $ (resp. in $\left( \Delta _{\frak{a}}\right) $);

$S\quad \left( S_{\frak{a}}\right) $ --- the set of simple roots (for
$\gf$ and $\af$ respectively);

$\alpha _{i}$ , $\left( \alpha _{\left( \frak{a}\right) j}\right) $ --- the $%
i$-th (resp. $j$-th) simple root for $\frak{g}$ $\left( \mathrm{{resp.}\frak{a%
}}\right) $; $i=0,\ldots ,r$,\ \ $\left( j=0,\ldots ,r_{\frak{a}}\right) $;

$\alpha _{i}^{\vee }$ , $\left( \alpha _{\left( \frak{a}\right) j}^{\vee
}\right) $-- the simple coroot for $\frak{g}$ $\left( \mathrm{{resp.}\frak{a}%
}\right) $ , $i=0,\ldots ,r$ ;\ \ $\left( j=0,\ldots ,r_{\frak{a}}\right) $;

$W$ , $\left( W_{\frak{a}}\right) $--- the Weyl group;

$C$ , $\left( C_{\frak{a}}\right) $--- the fundamental Weyl chamber;

$\bar{C}, \left(\bar{C_{\frak{a}}}\right)$ --- the closure of the
fundamental Weyl chamber;

$\epsilon \left( w\right) :=\left( -1\right) ^{\mathrm{length}(w)}$;

$\rho $\ , $\left( \rho _{\frak{a}}\right) $\ --- the Weyl vector;

$L^{\mu }$\ $\left( L_{\frak{a}}^{\nu }\right) $\ --- the integrable module
of $\frak{g}$ with the highest weight $\mu $\ ; (resp. integrable $\af$-module
with the highest weight $\nu $);

$\mathcal{N}^{\mu }$ , $\left( \mathcal{N}_{\frak{a}}^{\nu }\right) $ ---
the weight diagram of $L^{\mu }$ (resp. ${}L_{\frak{a}}^{\nu }$ );

$P$ (resp. $P_{\frak{a}} $) \ --- the weight lattice;

$P^{+}$ (resp. $P_{\frak{a}}^{+} $) \ --- the dominant weight lattice;

$m_{\xi }^{\left( \mu \right) }$ , $\left( m_{\zeta }^{\left( \nu \right)
}\right) $ --- the multiplicity of the weight $\xi \in P$ \ $\left( \mathrm{{%
resp. }\in P_{\frak{a}}}\right) $ in $L^{\mu }$, (resp. in $\zeta \in
L_{\frak{a}}^{\nu } $);

$ch\left( L^{\mu }\right) $ (resp. $\mathrm{ch}\left( L_{\frak{a}}^{\nu
}\right) $)--- the formal character of $L^{\mu }$ (resp. of $L_{\frak{a}}^{\nu
} $);

$ch\left( L^{\mu }\right) =\frac{\sum_{w\in W}\epsilon (w)e^{w\circ (\mu
+\rho )-\rho }}{\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{%
\mathrm{{mult}\left( \alpha \right) }}}$ --- the Weyl-Kac formula;

$R:=\prod_{\alpha \in \Delta ^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{{%
mult}\left( \alpha \right) }}\quad $ (resp. $R_{\frak{a}}:=\prod_{\alpha \in
\Delta _{\frak{a}}^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{mult}_{\frak{a}%
}\left( \alpha \right) } $)--- the Weyl denominator.

\section{Orthogonal subalgebra and\\ singular elements}

\label{sec:recurr-form-branch}

In this section we shall show how the recurrent approach to branching
problem leads naturally to a presentation of a formal character of
$\frak{g}$-module in terms characters corresponding to a set
of parabolic (generalized) Verma modules. Consider a reductive
Lie algebra $\frak{g}$ and its reductive subalgebra $\frak{a}\subset \frak{g}$.
Let $L^{\mu} $ be the highest weight integrable module of $\frak{g}$, $\mu \in
P^{+}$. Let $L^{\mu }$ be completely reducible with respect to $\frak{a}$,
\begin{equation*}
L_{\frak{g}\downarrow \frak{a}}^{\mu }=\bigoplus\limits_{\nu \in P_{\frak{a}%
}^{+}}b_{\nu }^{\left( \mu \right) }L_{\frak{a}}^{\nu }.
\end{equation*}
Using the projection operator $\pi _{\frak{a}}$ (to the weight space $\frak{%
h_{a}}^{\ast }$) one can write this decomposition in terms of formal
characters:
\begin{equation}
\pi _{\frak{a}}ch\left( L^{\mu }\right) =\sum_{\nu \in P_{\frak{a}%
}^{+}}b_{\nu }^{(\mu )}ch\left( L_{\frak{a}}^{\nu }\right) .
\label{branching1}
\end{equation}
The module $L^{\mu }$ has the BGG resolution (see \cite
{bernstein1976category,bernstein1975differential,bernstein1971structure} and
\cite{humphreys2008representations}). All the members of the filtration
sequence are the direct sums of Verma modules and all their highest weights $%
\nu $ are strongly linked to $\mu $:
\begin{equation*}
\left\{ \nu \right\} =\left\{ w\left( \mu +\rho \right) -\rho |w\in
W\right\} .
\end{equation*}

\subsection{Orthogonal subalgebra}

Let $\frak{h}_{\frak{a}}$ be a Cartan subalgebra of $\mathfrak{g}$. For $%
\mathfrak{a}\hookrightarrow \frak{g}$ introduce the ''orthogonal partner'' $%
\mathfrak{a}_{\bot }\hookrightarrow \frak{g}$ .

Consider the root subspace $\frak{h}_{\perp \frak{a}}^{\ast }$ orthogonal to
$\frak{a}$,
\begin{equation*}
\frak{h}_{\perp \frak{a}}^{\ast }:=\left\{ \eta \in \frak{h}^{\ast }|\forall
h\in \frak{h}_{\frak{a}};\eta \left( h\right) =0\right\} ,
\end{equation*}
and the roots (correspondingly -- positive roots) of $\frak{g}$ orthogonal
to $\frak{a}$,
\begin{eqnarray}
\Delta _{\frak{a}_{\perp }} &:&=\left\{ \beta \in \Delta _{\frak{g}}|\forall
h\in \frak{h}_{\frak{a}};\beta \left( h\right) =0\right\} ,
\label{delta a ort} \\
\Delta _{\frak{a}_{\perp }}^{+} &:&=\left\{ \beta ^{+}\in \Delta _{\frak{g}%
}^{+}|\forall h\in \frak{h}_{\frak{a}};\beta ^{+}\left( h\right) =0\right\} .
\notag
\end{eqnarray}
Let $W_{\frak{a}_{\perp }}$ be the subgroup of $W$ generated by the
reflections $w_{\beta }$ with the roots $\beta \in \Delta _{\frak{a}_{\perp
}}^{+}$. The subsystem $\Delta _{\frak{a}_{\perp }}$ determines the
subalgebra $\frak{a}_{\perp }$ with the Cartan subalgebra $\frak{h}_{\frak{a}%
_{\perp }}$. Let
\begin{equation*}
\frak{h}_{\perp }^{\ast }:=\left\{ \eta \in \frak{h}_{\perp \frak{a}}^{\ast
}|\forall h\in \frak{h}_{\frak{a}\oplus \frak{a}_{\perp }};\eta \left(
h\right) =0\right\}
\end{equation*}
so that $\frak{g}$ has the subalgebras
\begin{eqnarray}
\widetilde{\frak{a}_{\perp }} :=\frak{a}_{\perp }\oplus \frak{h}_{\perp }
\qquad
\widetilde{\frak{a}} :=\frak{a}\oplus \frak{h}_{\perp }.
\end{eqnarray}
Notice that $\mathfrak{a} \oplus \mathfrak{a}_{\bot}$ in general is not a
subalgebra in $\mathfrak{g}$.

For the Cartan subalgebras we have the decomposition
\begin{equation}
\frak{h}=\frak{\frak{h}_{\frak{a}}}\oplus \frak{h}_{\frak{a}_{\perp }}\oplus
\frak{h}_{\perp }=\frak{\frak{h}_{\widetilde{\frak{a}}}}\oplus \frak{h}_{%
\frak{a}_{\perp }}=\frak{\frak{h}_{\widetilde{\frak{a}_{\perp }}}}\oplus
\frak{h}_{\frak{a}}.
\end{equation}
For $\frak{a}$ and $\frak{a}_{\perp }$ consider the
corresponding Weyl vectors, $\rho _{\frak{a}}$ and $\rho _{\frak{a}_{\perp
}} $. Form the so called ''defects'' $\mathcal{D}_{\frak{a}}$ and $\mathcal{%
D}_{\frak{a}_{\perp }}$ of the injection:
\begin{equation}
\mathcal{D}_{\frak{a}}:=\rho _{\frak{a}}-\pi _{\frak{a}}\rho , \qquad
\mathcal{D}_{\frak{a}_{\perp }}:=\rho _{\frak{a}_{\perp }}-\pi _{\frak{a}%
_{\perp }}\rho .  \label{defect ort}
\end{equation}

For $\mu \in P^{+}$ consider the linked weights $\left\{
\left( w(\mu +\rho )-\rho \right) |w\in W\right\} $. Consider the projections to
$h_{\frak{a}_{\perp }}^{\ast }$ additionally shifted by the defect $-%
\mathcal{D}_{\frak{a}_{\perp }}$:
\begin{equation*}
\mu _{\frak{a}_{\perp }}\left( w\right) :=\pi _{\frak{a}_{\perp }}\left[
w(\mu +\rho )-\rho \right] -\mathcal{D}_{\frak{a}_{\perp }},\quad w\in W.
\end{equation*}
Among the weights $\left\{ \mu _{\frak{a}_{\perp
}}\left( w\right) |w\in W\right\} $ one can always choose those located in
the fundamental chamber $\overline{C_{\frak{a}_{\perp }}}$. Let $U$ be the
set of representatives $u$ for the classes $W/W_{\frak{a}_{\perp }}$ such
that

\begin{equation}
U:=\left\{ u\in W|\quad \mu _{\frak{a}_{\perp }}\left( u\right) \in
\overline{C_{\frak{a}_{\perp }}}\right\} \quad .  \label{U-def}
\end{equation}
Thus we can form the subsets:
\begin{equation}
\mu _{\widetilde{\mathfrak{a}}}\left( u\right) :=\pi _{\widetilde{%
\mathfrak{a}}}\left[ u(\mu +\rho )-\rho \right] +\mathcal{D}_{\frak{a}%
_{\perp }},\quad u\in U,  \label{mu-a}
\end{equation}
and
\begin{equation}
\mu _{\frak{a}_{\perp }}\left( u\right) :=\pi _{\frak{a}_{\perp }}\left[
u(\mu +\rho )-\rho \right] -\mathcal{D}_{\frak{a}_{\perp }},\quad u\in U.
\label{mu-a-tilda}
\end{equation}

Notice that the subalgebra $\mathfrak{a}_{\bot}$ is regular by definition
since it is built on a subset of roots of the algebra $\mathfrak{g}$.

For the modules we are interested in the Weyl-Kac formula for $\mathrm{ch}%
\left( L^{\mu }\right) $ can be written in terms of singular elements \cite
{humphreys1997introduction},
\begin{equation*}
\Psi ^{\left( \mu \right) }:=\sum\limits_{w\in W}\epsilon (w)e^{w(\mu +\rho
)-\rho },
\end{equation*}
namely,
\begin{equation}
\mathrm{ch}\left( L^{\mu }\right) =\frac{\Psi ^{\left( \mu \right) }}{\Psi
^{\left( 0\right) }}=\frac{\Psi ^{\left( \mu \right) }}{R}.
\label{Weyl-Kac2}
\end{equation}
The same is true for the submodules $\mathrm{ch}\left( L_{\frak{a}}^{\nu
}\right) $ in (\ref{branching1})
\begin{equation*}
\mathrm{ch}\left( L_{\frak{a}}^{\nu }\right) =\frac{\Psi _{\frak{a}}^{\left(
\nu \right) }}{\Psi _{\frak{a}}^{\left( 0\right) }}=\frac{\Psi _{\frak{a}%
}^{\left( \nu \right) }}{R_{\frak{a}}},
\end{equation*}
with
\begin{equation*}
\Psi _{\frak{a}}^{\left( \nu \right) }:=\sum\limits_{w\in W_{\frak{a}%
}}\epsilon (w)e^{w(\nu +\rho _{_{\frak{a}}})-\rho _{_{\frak{a}}}}.
\end{equation*}

Applying formula (\ref{Weyl-Kac2}) to the branching rule (\ref{branching1})
we get the relation connecting the singular elements $\Psi ^{\left( \mu
\right) }$ and $\Psi _{\frak{a}}^{\left( \nu \right) }$ :
\begin{eqnarray}
\pi _{\frak{a}}\left( \frac{\sum_{w \in W}\epsilon (w )e^{w (\mu +\rho
)-\rho }}{\prod_{\alpha \in \Delta ^{+}}(1-e^{-\alpha })^{\mathrm{mult}%
(\alpha )}}\right) &=&\sum_{\nu \in P_{\frak{a}}^{+}}b_{\nu }^{(\mu )}\frac{%
\sum_{w \in W_{\frak{a}}}\epsilon (w )e^{w (\nu +\rho _{\frak{a}})-\rho _{%
\frak{a}}}}{\prod_{\beta \in \Delta _{\frak{a}}^{+}}(1-e^{-\beta })^{\mathrm{%
mult}_{\frak{a}}(\beta )}},  \notag  \label{eq:4} \\
\pi _{\frak{a}}\left( \frac{\Psi ^{\left( \mu \right) }}{R}\right)
&=&\sum_{\nu \in P_{\frak{a}}^{+}}b_{\nu }^{(\mu )}\frac{\Psi _{\frak{a}%
}^{\left( \nu \right) }}{R_{\frak{a}}}.
\end{eqnarray}

\subsection{Decomposing the singular element.}

\label{subsec:decomp-sing-element}

Now we shall perform a decomposition of the singular element $\Psi ^{\left(
\mu \right) }$ in terms of singular elements of the orthogonal partner
modules:

\begin{lemma}
Let $\frak{a}_{\bot }$ be the orthogonal partner of a reductive subalgebra $%
\frak{a}\hookrightarrow \frak{g}$ with $\frak{h}=\frak{\frak{h}_{\frak{a}}}%
\oplus \frak{h}_{\frak{a}_{\perp }}\oplus \frak{h}_{\perp }$, $\widetilde{%
\frak{a}_{\perp }}=\frak{a}_{\perp }\oplus \frak{h}_{\perp }$ and $%
\widetilde{\frak{a}}=\frak{a}\oplus \frak{h}_{\perp }$.

$L^{\mu }$ be the highest weight integrable module with $\mu \in P^{+}$ and

$\Psi ^{\left( \mu \right) }$\ -- the singular element of $L^{\mu }$.

Then the element $\Psi ^{\left( \mu \right) }$ can be decomposed into the
sum over $u\in U$ (see (\ref{U-def})) of singular elements $\Psi _{\frak{a}%
_{\perp }}^{\mu _{\frak{a}_{\perp }}\left( u\right) }$ with the coefficients
$\epsilon (u)e^{\mu _{\widetilde{\mathfrak{a}}}\left( u\right) }$:
\begin{equation}
\Psi ^{\left( \mu \right) }=\sum_{u\in U}\;\epsilon (u)e^{\mu _{\widetilde{%
\mathfrak{a}}}\left( u\right) }\Psi _{\frak{a}_{\perp }}^{\mu _{\frak{a}%
_{\perp }}\left( u\right) }.  \label{sing decomp main}
\end{equation}
\label{Psi-decomp-lemma}
\end{lemma}

\begin{proof}
Let
\[
u(\mu +\rho )=\pi _{\left( \aft\right) }u(\mu +\rho )+\pi _{\left(
\frak{a}_{\perp }\right) }u(\mu +\rho )
\]
with $u\in U$. For any $v\in W_{\frak{a}_{\bot }}$ consider
the singular weight $vu(\mu +\rho )-\rho $ and perform the decomposition:
\begin{equation}
\begin{array}{lcl}
vu(\mu +\rho )-\rho  & = & \pi _{\left( \frak{a}\right) }\left( u(\mu +\rho
)\right) -\rho +\rho _{\frak{a}_{\perp }}
\\
&  & +\ v\left( \pi _{\left( \aft_{\perp }\right) }u(\mu
+\rho )-\rho _{\frak{a}_{\perp }}+\rho _{\frak{a}_{\perp }}\right) -\rho _{%
\frak{a}_{\perp }} .
\end{array}
\label{sing-decomp-1}
\end{equation}
Use the defect $\mathcal{D}_{\frak{a}_{\bot }}$ (\ref{defect ort}) to
simplify the first line in (\ref{sing-decomp-1}):
\[
\begin{array}{r}
\pi _{\left( \aft\right) }\left( u(\mu +\rho )\right) -\rho +\rho _{%
\frak{a}_{\perp }}= \\
\pi _{\left( \aft\right) }\left( u(\mu +\rho )\right) -\pi _{\aft%
}\rho -\pi _{\af_{\bot }}\rho +\rho _{\frak{a}_{\bot }}= \\
=\pi _{\left( \aft\right) }\left( u(\mu +\rho )-\rho \right) +\mathcal{D}%
_{\frak{a}_{\bot }},
\end{array}
\]
and the second one:
\[
\begin{array}{c}
v\left( \pi _{\left( \frak{a}_{\perp }\right) }u(\mu +\rho
)-\rho _{\frak{a}_{\perp }}+\rho _{\frak{a}_{\perp }}\right) -\rho _{\frak{a}%
_{\perp }}= \\
v\left( \pi _{\left( \frak{a}_{\bot }\right) }u(\mu +\rho )-%
\mathcal{D}_{\frak{a}_{\bot }}-\pi _{\left( \frak{a}_{\bot }\right) }\rho
+\rho _{\frak{a}_{\bot }}\right)
-\rho _{\frak{a}_{\bot }}= \\
=v\left( \pi _{\left( \frak{a}_{\bot }\right) }\left[ u(\mu
+\rho )-\rho \right] -\mathcal{D}_{\frak{a}_{\bot }}+\rho _{\frak{a}_{\bot
}}\right) -\rho _{\frak{a}_{\bot }}.
\end{array}
\]
This provides the desired decomposition of the singular element $%
\Psi ^{\mu }$ in terms of singular elements $\Psi _{\frak{a}%
_{\perp }}^{\eta }$ of the $\frak{a}_{\perp }$-modules $L_{%
\frak{a}_{\perp }}^{\eta }$:
\begin{equation}
\begin{array}{l}
\Psi ^{\mu }=\sum_{u\in U}\sum_{v\in W_{\frak{a}_{\perp }}}\epsilon
(v)\epsilon (u)e^{vu(\mu +\rho )-\rho }= \\
=\sum_{u\in U}\epsilon (u)e^{\pi _{\aft}\left[ u(\mu +\rho )-\rho \right]
+\mathcal{D}_{\frak{a}_{\perp }}}\sum_{v\in W_{\frak{a}_{\perp }}}\epsilon
(v)e^{v\left( \pi _{\left( \frak{a}_{\perp }\right) }\left[
u(\mu +\rho )-\rho \right] -\mathcal{D}_{\frak{a}_{\perp }}+\rho _{\frak{a}%
_{\perp }}\right) -\rho _{\frak{a}_{\perp }}}= \\
=\sum_{u\in U}\;\epsilon (u)\Psi _{\af_{\perp }}^{\pi
_{\left( \frak{a}_{\perp }\right) }\left[ u(\mu +\rho )-\rho
\right] -\mathcal{D}_{\frak{a}_{\perp }}}e^{\pi _{\left( \aft\right) }%
\left[ u(\mu +\rho )-\rho \right] +\mathcal{D}_{\frak{a}_{\perp }}}.
\end{array}
\label{singular main}
\end{equation}
\end{proof}

\bigskip

\begin{remark}
This relation can be considered as a generalized form of the Weyl formula
for the singular element $\Psi _{\frak{g}}^{\mu }$: the vectors $\mu _{%
\widetilde{\mathfrak{a}}}\left( u\right) $ play the role of singular weights
while the alternating factors $\epsilon (u)$ are extended to $\epsilon
(u)\Psi _{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp }}\left( u\right) }$. In
fact when $\frak{a=g}$ both $\frak{a}_{\perp }$and $\frak{h}_{\perp }$are
zeros, $U=W$, and\ the original Weyl formula is reobtained so far as the singular elements $\epsilon (u)\Psi _{\frak{a}_{\perp
}}^{\mu _{\frak{a}_{\perp }}\left( u\right) }=\epsilon (u)$ become trivial.
 In the opposite
limit when $\frak{a=0}$, $\Delta _{\frak{a}_{\perp }}=\Delta _{\frak{g}}$, $%
\frak{h}_{\perp }^{\ast }=0$, $\frak{a}_{\perp }=\frak{g}$, $\mathcal{D}_{%
\frak{a}_{\perp }}=0$ and $U=W/W_{\frak{a}_{\perp }}=e$ the singular element
 $\Psi ^{\mu }$ is again reobtained, now via the trivilization of the set of
 vectors $\mu _{\af}\left( e\right) =0$.
\end{remark}

\begin{remark}

In \cite
{2010arXiv1007.0318L} the decomposition analogous to (\ref{singular main}) was
used to construct the recurrent relations for branching coefficients $k_{\xi
}^{\left( \mu \right) }$ corresponding to the injection $\frak{a}%
\hookrightarrow \frak{g}$:
\begin{equation}
\begin{array}{c}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{u\in U}\epsilon (u)\;\dim \left( L_{\frak{a}_{\perp }}^{\mu _{\frak{a}%
_{\perp }}\left( u\right) }\right) \delta _{\xi -\gamma _{0},\pi _{%
\widetilde{\frak{a}}}(u(\mu +\rho )-\rho )}+\right.  \\
\left. +\sum_{\gamma \in \Gamma _{\widetilde{\frak{a}}\rightarrow \frak{g}%
}}s\left( \gamma +\gamma _{0}\right) k_{\xi +\gamma }^{\left( \mu \right)
}\right) .
\end{array}
\label{recurrent rel}
\end{equation}
The recursion is goverened by the set $\Gamma _{\widetilde{\frak{a}}%
\rightarrow \frak{g}}$ called the injection fan. The latter is defined by the
carrier set $\left\{ \xi \right\} _{\frak{a}\rightarrow \frak{g}}$ for the
coefficient function $s(\xi )$
\begin{equation*}
\left\{ \xi \right\} _{\widetilde{\frak{a}}\rightarrow \frak{g}}:=\left\{
\xi \in P_{\widetilde{\frak{a}}}|s(\xi )\neq 0\right\}
\end{equation*}
appearing in the expansion
\begin{equation}
\prod_{\alpha \in \Delta ^{+}\setminus \Delta _{\bot }^{+}}\left( 1-e^{-\pi
_{\widetilde{\frak{a}}}\alpha }\right) ^{\mathrm{mult}(\alpha )-\mathrm{mult}%
_{\frak{a}}(\pi _{\widetilde{\frak{a}}}\alpha )}=-\sum_{\gamma \in P_{%
\widetilde{\frak{a}}}}s(\gamma )e^{-\gamma };\quad
\end{equation}
The weights in $\left\{ \xi \right\} _{\widetilde{\frak{a}}\rightarrow \frak{%
g}}$ are to be shifted by $\gamma _{0}$ -- the lowest vector in $\left\{ \xi
\right\} $ -- and the zero element is to be eliminated:
\begin{equation}
\Gamma _{\widetilde{\frak{a}}\rightarrow \frak{g}}=\left\{ \xi -\gamma
_{0}|\xi \in \left\{ \xi \right\} \right\} \setminus \left\{ 0\right\} .
\end{equation}

The recursion relation (\ref{recurrent rel}) was originally used to
describe branchings for integrable modules. Notice that there exists an
important class of modules that also can be reduced with the help of the
injection fan --
these are Verma modules.
\end{remark}

\subsection{Weyl-Verma formulas.}

\begin{statement}
\bigskip For an orthogonal subalgebra $\frak{a}_{\perp }$ in $\frak{g}$ (an
orthogonal partner of a reductive $\frak{a}\hookrightarrow \frak{g}$) the
character of an integrable highest weight module $L^{\mu }$ can be presented
as a combination (with integral coefficients) of parabolic Verma modules
distributed by the set of weights $e^{\mu _{\widetilde{\mathfrak{a}}}\left(
u\right) }$:
\begin{equation}
\mathrm{ch}\left( L^{\mu }\right) =\sum_{u\in U}\;\epsilon (u)e^{\mu _{%
\widetilde{\frak{a}}}\left( u\right) }\mathrm{ch}M_{I}^{\mu _{\frak{a}%
_{\perp }}\left( u\right) },  \label{gen Weyl-Verma}
\end{equation}
where $U:=\left\{ u\in W|\quad \mu _{\frak{a}_{\perp }}\left( u\right) \in
\overline{C_{\frak{a}_{\perp }}}\right\} $ and $I$ is such a subset of $S$ that
$\Delta _{I}^{+}$ is
equivalent to $\Delta _{\frak{a}_{\perp }}^{+}$.
\end{statement}

\bigskip
\begin{proof}
By the definition (\ref{delta a ort}) the subalgebra $%
\mathfrak{a}_{\bot }$ is regular and reductive. Consider its Weyl
denominator $R_{\frak{a}_{\perp }}:=\prod_{\alpha \in \Delta _{\frak{a}%
_{\perp }}^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{mult}_{\frak{a}}%
\mathrm{\left( \alpha \right) }}$ and the element $R_{J}:=\prod_{\alpha \in
\Delta ^{+}\setminus \Delta _{\frak{a}_{\perp }}^{+}}\left( 1-e^{-\alpha
}\right) ^{\mathrm{mult}(\alpha )}$ as the factors in $R$: $\quad $%
\begin{equation*}
R=R_{J}R_{\frak{a}_{\perp }}.
\end{equation*}
According to this factorization and the decomposition (\ref{sing decomp main}%
) the character $\mathrm{ch}\left( L^{\mu }\right) $ can be written as
\begin{eqnarray*}
\mathrm{ch}\left( L^{\mu }\right) &=&\left( R_{J}\right) ^{-1}\left( R_{%
\frak{a}_{\perp }}\right) ^{-1}\Psi ^{\mu }=\left( R_{J}\right)
^{-1}\sum_{u\in U}\;e^{\mu _{\widetilde{\frak{a}}}\left( u\right) }\epsilon
(u)\left( R_{\frak{a}_{\perp }}\right) ^{-1}\Psi _{\frak{a}_{\perp }}^{\mu _{%
\frak{a}_{\perp }}\left( u\right) } \\
&=&\left( R_{J}\right) ^{-1}\sum_{u\in U}\;e^{\mu _{\widetilde{\frak{a}}%
}\left( u\right) }\epsilon (u)\mathrm{ch}\left( L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp
}}\left( u\right) }\right),
\end{eqnarray*}
where $\left\{ L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp }}\left(
u\right) }|u\in U\right\} $ is the set of finite-dimensional $\frak{a}%
_{\perp }$-modules with the highest weights $\mu _{\frak{a}_{\perp }}\left(
u\right) $. We are interested in nontrivial subalgebras $\frak{a}$ and
correspondingly in nontrivial $\frak{a}_{\perp }$ (the case of a trivial
orthogonal subalgebra was considered above (see Remark 1)). This means that $r_{\frak{a}}\geq 1$ and $r_{\frak{a}%
_{\perp }}<r$. Due to the fact that any maximal regular subalgebra has the
Dynkin scheme obtained by one or two node subtractions from the extended
Dynkin scheme and the extended scheme has at most one dependent root (the
highest root) the set of roots $\Delta _{\frak{a}_{\perp }}^{+}$ is always
equivalent to the one $\Delta _{I}^{+}$ generated\ by some subset $I\subset
S $ of simple roots.

>>>>

\title{{\bf Affine.m} -- {\it Mathematica} package for Lie algebras of finite, affine and extended affine type}
\author{A A Nazarov$^{1,2}$\\
{\small $^1$ Theoretical Department, SPb State University}\\
{\small 198904, Sankt-Petersburg, Russia}\\
{\small$^{2}$ Chebyshev Laboratory,}\\
{\small Department of Mathematics and Mechanics, SPb State University}\\
{\small 199178, Saint-Petersburg, Russia}\\
{\small email: antonnaz@gmail.com}}

@@

\begin{abstract}
  We present {\it Mathematica} package for computations in the representation theory of Lie algebras of finite, affine and extended affine types. 
\end{abstract}

\section{Introduction}
\end{document}
